---
title: "Model comparison (generated data)"
output:
  pdf_document: default
  html_document: default
date: "2023-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr)
library(DescTools)
library(scales)
library(ppcor)
library(ggdag)
library(brms)
library(tidybayes)
library(marginaleffects)

theme_set(theme_bw())
```

```{r}
packageVersion("brms")
```


### Loading the data 

Loading the last step ($t_300$) and transforming dependent variables to z-scores:

```{r}
last_step <- read.csv('data/model-N570_S1000_p50000.csv') %>%
  filter(extinct != 1) %>% # remove extinct words
  group_by(step) %>% # group by time-step
  mutate(relative_frequency = count / sum(count)) %>% 
  filter(step == 300-1 & extinct != 1)  %>% 
  mutate(z_age = (age - mean(age))/sd(age),
         z_freq = (log(relative_frequency) - 
                     mean(log(relative_frequency)))/sd(log(relative_frequency)))
```

```{r}
head(last_step)
```


Plotting the distribution of number of meanings:

```{r}
last_step %>%
  ggplot(aes(x=meanings)) +
  geom_bar(position = 'identity', fill='blue', alpha = 0.5)
```


### Analysis

#### Correlations

Spearman correlation between age and frequency:

```{r}
cor.test(last_step$z_age, last_step$z_freq, method = 'spearman', exact=FALSE)
```
#### Model selection 

```{r}
## Zero-inflated
mod1 <- 
  brm(meanings - 1 ~ 1, 
      data = last_step,
      family = zero_inflated_poisson())

## Standard poisson 
mod2 <- 
  brm(meanings - 1 ~ 1,
      data = last_step, 
      family = poisson())

## Poisson lognormal
mod3 <- 
  brm(meanings - 1 ~ 1 + (1 | id), 
      data = last_step,
      family = poisson(), 
      save_pars = save_pars(all = TRUE))
```

```{r}
mod1 <- add_criterion(mod1, criterion = "loo")
mod2 <- add_criterion(mod2, criterion = "loo")
mod3 <- add_criterion(
  mod3, criterion = "loo", moment_matching = T
)

loo_compare(mod1, mod2, mod3) %>% 
  print(simplify = F)
```


```{r}
a <- pp_check(mod1, ndraws = 100, type = "bars",) +
  xlim(-1, 10) +
  ggtitle('Zero-inflated')
b <- pp_check(mod2, ndraws = 100, type = "bars",) +
  xlim(-1, 10) +
  ggtitle('Standard poisson')
c <- pp_check(mod3, ndraws = 100, type = "bars",)+
  xlim(-1, 10) +
  ggtitle('Poisson lognormal')

ggarrange(a, b, c,
          common.legend = TRUE, 
          legend = 'right')
```

```{r}
rbind(bayes_R2(mod1), 
      bayes_R2(mod2), 
      bayes_R2(mod3)) %>%
  as_tibble() %>%
  mutate(model = c("mod1", "mod2", "mod3"),
         r_square_posterior_mean = round(Estimate, digits = 6)) %>%
  dplyr::select(model, r_square_posterior_mean)
```

The Poisson with random effects per observation seems weird, since the fit is already too good for the intercept only model -- possibility of overfitting? Let's use the zero-inflated model since it's posterior predictive checks look better, and it has the second best loo. 

#### Final analysis 

Fitting the 4 possible models:

- intercept only
- frequency only 
- longevity only 
- full model 

```{r}
## intercept only
mod3.1 <- 
  brm(meanings - 1 ~ 1,
      data = last_step,
      family = zero_inflated_poisson())

## frequency only
mod3.2 <- 
  brm(meanings - 1 ~ 1 + z_freq, 
      data = last_step,
      family = zero_inflated_poisson())

## age only
mod3.3 <- 
  brm(meanings - 1 ~ 1 + z_age, 
      data = last_step,
      family = zero_inflated_poisson())

## full model
mod3.4 <- 
  brm(meanings - 1 ~ 1 + z_age + z_freq, 
      data = last_step,
      family = zero_inflated_poisson())
```

```{r}
mod3.1 <- add_criterion(
  mod3.1, criterion = "waic"
)

mod3.2 <- add_criterion(
  mod3.2, criterion = "waic"
)

mod3.3 <- add_criterion(
  mod3.3, criterion = "waic"
)

mod3.4 <- add_criterion(
  mod3.4, criterion = "waic"
)
```


WAIC comparison:

```{r}
loo_compare(mod3.1, mod3.2, mod3.3, mod3.4, criterion = 'waic') %>% 
  print(simplify = F)
```

Getting coefficients for plotting: 

```{r}
full <- avg_comparisons(mod3.4) %>%
  posterior_draws(shape = "rvar") %>%
  ggplot(aes(y = term, xdist = rvar)) + 
  stat_slabinterval() +
  xlim(0, 1) +
  ggtitle('Full model') +
  ylab('') + 
  xlab('') +
  geom_vline(xintercept = 0, 
             linetype="dotted",
             color = 'red')

only_freq <- avg_comparisons(mod3.2) %>%
  posterior_draws(shape = "rvar") %>%
  ggplot(aes(y = term, xdist = rvar)) + 
  stat_slabinterval() +
  xlim(0, 1) +
  ggtitle('Only frequency model') +
  ylab('') + 
  xlab('') +
  geom_vline(xintercept = 0, 
             linetype="dotted",
             color = 'red')

only_age <- avg_comparisons(mod3.3) %>%
  posterior_draws(shape = "rvar") %>%
  ggplot(aes(y = term, xdist = rvar)) + 
  stat_slabinterval() +
  xlim(0, 1) +
  ggtitle('Only longevity model') +
  ylab('') + 
  xlab('') +
  geom_vline(xintercept = 0, 
             linetype="dotted",
             color = 'red')
```

Coefficients comparison:

```{r}
ggarrange(full, only_freq, only_age, nrow=3, ncol=1)
```

Predictions:

```{r}
longevity <- plot_predictions(
    mod3.4,
    condition = "z_age") +
    labs(y = "meanings") +
  geom_count(data = last_step, 
             aes(x = z_age, y = meanings - 1),
             alpha = 0.2, 
             color = 'blue') +
  xlab('Longevity (z-scored)') +
  ylab('Number of additional meanings') +
  scale_size_area()

frequency <- plot_predictions(
    mod3.4,
    condition = "z_freq") +
    labs(y = "meanings") +
  geom_count(data = last_step, 
             aes(x = z_freq, y = meanings - 1),
             alpha = 0.2,
             color = 'blue') +
  xlab('Log relative frequency (z-scored)') +
  ylab('Number of additional meanings') +
  scale_size_area()

ggarrange(longevity, frequency, ncol = 2, nrow = 1, labels = c("A", "B"))
```


Posterior predictive check: 

```{r}
pp_check(mod3.4, ndraws = 100, type = "bars") +
  xlim(-1, 10) +
  ylim(0, 2500) +
  ggtitle('Zero-inflated')
```
